### 2024-01-18 outputs/paper/paper.qmd
updated paper and updated data cleaning

---
title: "Is Crime Really On A Rise In Toronto? An Analysis of The Past And Current Crime Rates of Toronto"
author: 
  - Adrian Ly
thanks: "Code and data are available at: LINK."
date: today
date-format: long
abstract: "Current mainstream news outlets are pushing a narrative that crime in Toronto is on a rise. "
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(knitr)
library(janitor)
library(lubridate)
library(opendatatoronto)
library(tidyverse)
library(dplyr)
```

# Introduction

You can and should cross-reference sections and sub-sections.

The remainder of this paper is structured as follows. @sec-data....

# Data {#sec-data}

The data utilized throughout this paper was obtained through the City of Toronto's

Some of our data is of penguins (@fig-bills), from @palmerpenguins.

```{r}
#| label: fig-bills
#| fig-cap: Bills of penguins
#| echo: false

ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = "none") +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

```{r}
#| label: fig-planes
#| fig-cap: Relationship between wing length and width
#| echo: false
#| warning: false
#| message: false

analysis_data <- read_csv(here::here("outputs/data/analysis_data.csv"))

analysis_data |> 
  ggplot(aes(x = width, y = length)) +
  geom_point(alpha = 0.8) +
  theme_minimal() +
  labs(x = "Wing width (mm)",
       y = "Wing length (mm)")
```

Talk way more about it.

# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

$$
\text{Percentage Change} = \left( \frac{\text{New Value} - \text{Old Value}}{\text{Old Value}} \right) \times 100
$$

```{=tex}
\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}
```
We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

Our results are summarized in @fig-avgcrimerates.

```{r}
#| label: fig-avgcrimerates
#| fig-cap: Average Crime Rates
#| echo: false

# Reading the data

average_crime_rates_per_year <- read_csv("outputs/data/average_crime_rates_data.csv")

# Creating the line graph for average crime rates
ggplot(average_crime_rates_per_year, aes(x = year, y = average_rate, color = crime)) +
  geom_line() +
  labs(title = "Average Crime Rates by Type Across Years (2014-2023)",
       x = "Year",
       y = "Average Crime Rate",
       color = "Crime Type") +
  theme_minimal()
```

```{r}
#| label: fig-percentagecrimerates
#| fig-cap: Average Crime Rates
#| echo: false

# Reading the data
percentage_change_data <- read_csv("outputs/data/percentage_change_data.csv")

# Creating the line graph for percentage changes
ggplot(percentage_change_data, aes(x = year, y = percentage_change, color = crime)) +
  geom_line() +
  labs(title = "Yearly Percentage Change in Average Crime Rates by Type (2014-2023)",
       x = "Year",
       y = "Percentage Change in Average Rate",
       color = "Crime Type") +
  theme_minimal()
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-percentagecrimerates20142023
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

# Reading the data
final_percentage_change_table <- read.csv("outputs/data/percentage_change_table_2014_2023.csv")
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-percentagecrimerates20222023
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

percentage_change_table <- read.csv("outputs/data/percentage_change_table_2022_2023.csv")
```

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {.unnumbered}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

pp_check(first_model) +
  theme_classic() +
  theme(legend.position = "bottom")

posterior_vs_prior(first_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

plot(first_model, "trace")

plot(first_model, "rhat")
```

\newpage

# References